---
title: "iSEEtree: interactive visualisation for microbiome data"
author: 
  - name: Giulio Benedetti
    affiliation: University of Turku
    email: giulio.benedetti@utu.fi
output:
    BiocStyle::html_document:
        fig_height: 7
        fig_width: 10
        toc: true
        toc_float: true
        toc_depth: 2
        number_sections: true
        self_contained: true
        code_folding: show
date: "`r doc_date()`"
package: "`r pkg_ver('iSEEtree')`"
vignette: >
    %\VignetteIndexEntry{iSEEtree}
    %\VignetteEngine{knitr::rmarkdown}
    %\VignetteEncoding{UTF-8}
---

```{r, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r vignetteSetup, echo=FALSE, message=FALSE, warning=FALSE}
## Bib setup
library(RefManageR)

## Write bibliography information
bib <- c(
    R = citation(),
    BiocStyle = citation("BiocStyle")[1],
    knitr = citation("knitr")[1],
    RefManageR = citation("RefManageR")[1],
    rmarkdown = citation("rmarkdown")[1],
    testthat = citation("testthat")[1]
)
```

# Introduction

## Motivation

iSEEtree is a Bioconductor package for the interactive visualisation of
microbiome data stored in a TreeSummarizedExperiment (TreeSE) container. On the
one side, it leverages and extends the graphics of the iSEE package, which is
designed for the generic SummarizedExperiment class. On the other side, it
employs the statistical and visual tools for microbiome data science provided by
the mia family of packages. Thus, iSEE and mia represent the two building blocks
of iSEEtree. Detailed introductory material on these two frameworks is available
in the [iSEE-verse website](https://isee.github.io/) and the
[OMA Bioconductor book](https://microbiome.github.io/OMA/docs/devel/),
respectively.

iSEEtree is meant for new and experienced users alike, who desire to create and
interact with several graphics for microbiome data, without the need for an
in-depth knowledge of the underlying mia functionality. Current
microbiome-specific panels include phylogenetic trees, ordination plots and compositional plots, which can be further explored below in this article. Other
more generic panels are also reused from the iSEE package and can be
experimented in [this article](https://isee.github.io/iSEE/articles/basic.html).

## Panels

iSEEtree derives its microbiome-related visualisation methods from the
[miaViz](https://microbiome.github.io/miaViz/) package, which is code-based and
requires basic knowledge of R programming and microbiome data structures. The
following panels represent an easy-to-use interactive version of the miaViz
plotting functions:

- [AbundanceDensityPlot](https://microbiome.github.io/iSEEtree/reference/AbundanceDensityPlot.html):
  a density plot of the top features, where every line is a feature and the x
  axis shows its abundance for different samples. Its interpretation is
  explained in the OMA chapter on
  [Exploration](https://microbiome.github.io/OMA/docs/devel/pages/12_quality_control.html).
- [AbundancePlot](https://microbiome.github.io/iSEEtree/reference/AbundancePlot.html):
  a compositional barplot of the samples, where every bar is a sample composed
  by different features in different colours. Its interpretation is explained
  in the OMA chapter on
  [Community Composition](https://microbiome.github.io/OMA/docs/devel/pages/21_microbiome_community.html).
- [RDAPlot](https://microbiome.github.io/iSEEtree/reference/RDAPlot.html): an
  supervised ordination plot of the samples, where every dot is a sample on a
  reduced dimensional space and every arrow reflects the contribution of a
  sample variable. Its interpretation is explained in the OMA chapter on
  [Community Similarity](https://microbiome.github.io/OMA/docs/devel/pages/20_beta_diversity.html).
- [RowTreePlot](https://microbiome.github.io/iSEEtree/reference/RowTreePlot.html):
  a phylogenetic tree of the features, where every tip is a feature and two
  neighbouring tips are more closely related than two far apart from each other.

By default, the iSEEtree layout also includes the following panels inherited by
iSEE:

- [RowDataTable](https://isee.github.io/iSEE/articles/basic.html#row-data-tables)
- [ColumnDataTable](https://isee.github.io/iSEE/articles/basic.html#column-data-tables)
- [ReducedDimensionPlot](https://isee.github.io/iSEE/articles/basic.html#reduced-dimension-plots)
- [ComplexHeatmapPlot](https://isee.github.io/iSEE/articles/basic.html#heat-maps)

The [ColumnDataPlot](https://isee.github.io/iSEE/articles/basic.html#coldataplot)
could also prove useful for the visualisation of column variables such as
alpha diversity indices. Its interpretation is explained in the OMA chapter on
[Community Diversity](https://microbiome.github.io/OMA/docs/devel/pages/14_alpha_diversity.html).

# Tutorial

## Installation
R is an open-source statistical environment which can be easily modified to
enhance its functionality via packages. iSEEtree is an R package available on
[Bioconductor](https://bioconductor.org/). R can be installed on any operating
system from [CRAN](https://cran.r-project.org/) after which you can install
iSEEtree by using the following commands in your R session:

```{r install, eval=FALSE}
if (!requireNamespace("BiocManager", quietly = TRUE))
    install.packages("BiocManager")

BiocManager::install("iSEEtree")
```

## Example

The panels described above can be generated for a model TreeSE object in the
following example:

```{r start, message=FALSE, warning=FALSE}
library(iSEEtree)
library(mia)
library(scater)

# Import TreeSE
data("Tengeler2020", package = "mia")
tse <- Tengeler2020

# Add relabundance assay
tse <- transformAssay(tse, method = "relabundance")

# Add reduced dimensions
tse <- runMDS(tse, assay.type = "relabundance")

# Launch iSEE
if (interactive()) {
  iSEE(tse)
}
```

```{r screenfun, eval=!exists("SCREENSHOT"), include=FALSE}
SCREENSHOT <- function(x, ...) knitr::include_graphics(x)
```

```{r screenplot, echo=FALSE, out.width="100%"}
SCREENSHOT("screenshots/get_started.png", delay=20)
```

# Resources

## Citation

We hope that iSEEtree will be useful for your research. Please use the following information to cite the package and the overall approach. Thank you!

```{r citation}
## Citation info
citation("iSEEtree")
```

## Background Knowledge

iSEEtree is based on many other packages and in particular on those that have implemented the infrastructure needed for dealing with omics data, microbiome
data and interactive visualisation. That is, packages like
[_SummarizedExperiment_](https://bioconductor.org/packages/3.18/bioc/html/SummarizedExperiment.html),
[_TreeSummarizedExperiment_](https://bioconductor.org/packages/3.18/bioc/html/TreeSummarizedExperiment.html),
[_mia_](https://bioconductor.org/packages/3.18/bioc/html/mia.html),
[_iSEE_](https://bioconductor.org/packages/3.18/bioc/html/iSEE.html) and
[_shiny_](https://cran.r-project.org/web/packages/shiny/).

If you are asking yourself the question “Where do I start using Bioconductor?”
you might be interested in
[this blog post](https://lcolladotor.github.io/2014/10/16/startbioc/#.VkOKbq6rRuU).

## Help

As package developers, we try to explain clearly how to use our packages and in
which order to use the functions. But R and Bioconductor have a steep learning
curve so it is critical to learn where to ask for help. The blog post quoted
above mentions some but we would like to highlight the
[Bioconductor support site](https://support.bioconductor.org/) as the main
resource for getting help: remember to use the iSEEtree tag and check the
[older posts](https://support.bioconductor.org/t/iSEEtree/). Other alternatives
are available such as creating GitHub issues and tweeting. However, please note
that if you want to receive help you should adhere to the
[posting guidelines](https://www.bioconductor.org/help/support/posting-guide/).
It is particularly critical that you provide a small reproducible example and
your session information so package developers can track down the source of the
error.

# Reproducibility

iSEEtree was made possible thanks to:

- R `r Citep(bib[["R"]])`
- `r Biocpkg("BiocStyle")` `r Citep(bib[["BiocStyle"]])`
- `r CRANpkg("knitr")` `r Citep(bib[["knitr"]])`
- `r CRANpkg("RefManageR")` `r Citep(bib[["RefManageR"]])`
- `r CRANpkg("rmarkdown")` `r Citep(bib[["rmarkdown"]])`
- `r CRANpkg("testthat")` `r Citep(bib[["testthat"]])`

This package was developed using
[_usethis_](https://cran.r-project.org/web/packages/usethis/).

R session information:

```{r reproduce, echo=FALSE}
## Session info
options(width = 120)
sessionInfo()
```

# References

This vignette was generated using `r Biocpkg("BiocStyle")` `r Citep(bib[["BiocStyle"]])`
with `r CRANpkg("knitr")` `r Citep(bib[["knitr"]])` and `r CRANpkg("rmarkdown")` `r Citep(bib[["rmarkdown"]])` running behind the scenes. Citations were generated with `r CRANpkg("RefManageR")` `r Citep(bib[["RefManageR"]])`.

```{r vignetteBiblio, results="asis", echo=FALSE, warning=FALSE, message=FALSE}
## Print bibliography
PrintBibliography(bib, .opts = list(hyperlink = "to.doc", style = "html"))
```
