---
title: "iSEEtree: interactive visualisation for microbiome data"
author: 
  - name: Giulio Benedetti
    affiliation: University of Turku
    email: giulio.benedetti@utu.fi
output:
    BiocStyle::html_document:
        fig_height: 7
        fig_width: 10
        toc: true
        toc_float: true
        toc_depth: 2
        number_sections: true
        self_contained: true
        code_folding: show
date: "`r doc_date()`"
package: "`r pkg_ver('iSEEtree')`"
vignette: >
    %\VignetteIndexEntry{iSEEtree}
    %\VignetteEngine{knitr::rmarkdown}
    %\VignetteEncoding{UTF-8}
---

```{r, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r vignetteSetup, echo=FALSE, message=FALSE, warning=FALSE}
## Bib setup
library(RefManageR)

## Write bibliography information
bib <- c(
    R = citation(),
    BiocStyle = citation("BiocStyle")[1],
    knitr = citation("knitr")[1],
    RefManageR = citation("RefManageR")[1],
    rmarkdown = citation("rmarkdown")[1],
    testthat = citation("testthat")[1]
)
```

# Basics

## Install iSEEtree
R is an open-source statistical environment which can be easily modified to
enhance its functionality via packages. iSEEtree is an R package available on
[Bioconductor](https://bioconductor.org/). R can be installed on any operating
system from [CRAN](https://cran.r-project.org/) after which you can install
iSEEtree by using the following commands in your R session:

```{r install, eval=FALSE}
if (!requireNamespace("BiocManager", quietly = TRUE))
    install.packages("BiocManager")

BiocManager::install("iSEEtree")
```

## Required knowledge
iSEEtree is based on many other packages and in particular on those that have implemented the infrastructure needed for dealing with omics data, microbiome
data and interactive visualisation. That is, packages like
[_SummarizedExperiment_](https://bioconductor.org/packages/3.18/bioc/html/SummarizedExperiment.html),
[_TreeSummarizedExperiment_](https://bioconductor.org/packages/3.18/bioc/html/TreeSummarizedExperiment.html),
[_mia_](https://bioconductor.org/packages/3.18/bioc/html/mia.html),
[_iSEE_](https://bioconductor.org/packages/3.18/bioc/html/iSEE.html) and
[_shiny_](https://cran.r-project.org/web/packages/shiny/).

If you are asking yourself the question “Where do I start using Bioconductor?”
you might be interested in
[this blog post](https://lcolladotor.github.io/2014/10/16/startbioc/#.VkOKbq6rRuU).

## Asking for help
As package developers, we try to explain clearly how to use our packages and in
which order to use the functions. But R and Bioconductor have a steep learning
curve so it is critical to learn where to ask for help. The blog post quoted
above mentions some but we would like to highlight the
[Bioconductor support site](https://support.bioconductor.org/) as the main
resource for getting help: remember to use the iSEEtree tag and check the
[older posts](https://support.bioconductor.org/t/iSEEtree/). Other alternatives
are available such as creating GitHub issues and tweeting. However, please note
that if you want to receive help you should adhere to the
[posting guidelines](https://www.bioconductor.org/help/support/posting-guide/).
It is particularly critical that you provide a small reproducible example and
your session information so package developers can track down the source of the
error.

## Citing iSEEtree
We hope that iSEEtree will be useful for your research. Please use the following information to cite the package and the overall approach. Thank you!

```{r citation}
## Citation info
citation("iSEEtree")
```

# Quick start to iSEEtree

```{r start, message=FALSE, warning=FALSE}
library(iSEEtree)
library(mia)
library(scater)

# Import TreeSE
data("Tengeler2020", package = "mia")
tse <- Tengeler2020

# Add relabundance assay
tse <- transformAssay(tse, method = "relabundance")

# Add reduced dimensions
tse <- runMDS(tse, assay.type = "relabundance")

# Launch iSEE
if (interactive()) {
  iSEE(tse)
}
```

```{r screenfun, eval=!exists("SCREENSHOT"), include=FALSE}
SCREENSHOT <- function(x, ...) knitr::include_graphics(x)
```

```{r screenplot, echo=FALSE, out.width="100%"}
SCREENSHOT("screenshots/get_started.png", delay=20)
```

# Reproducibility
iSEEtree was made possible thanks to:

- R `r Citep(bib[["R"]])`
- `r Biocpkg("BiocStyle")` `r Citep(bib[["BiocStyle"]])`
- `r CRANpkg("knitr")` `r Citep(bib[["knitr"]])`
- `r CRANpkg("RefManageR")` `r Citep(bib[["RefManageR"]])`
- `r CRANpkg("rmarkdown")` `r Citep(bib[["rmarkdown"]])`
- `r CRANpkg("testthat")` `r Citep(bib[["testthat"]])`

This package was developed using
[_usethis_](https://cran.r-project.org/web/packages/usethis/).

Code for creating the vignette.

```{r createVignette, eval=FALSE}
## Create the vignette
library(rmarkdown)
render("iSEEtree.Rmd", "BiocStyle::html_document")

## Extract the R code
library(knitr)
knit("iSEEtree.Rmd", tangle = TRUE)
```

R session information.

```{r reproduce3, echo=FALSE}
## Session info
options(width = 120)
sessionInfo()
```

# Bibliography

This vignette was generated using `r Biocpkg("BiocStyle")` `r Citep(bib[["BiocStyle"]])`
with `r CRANpkg("knitr")` `r Citep(bib[["knitr"]])` and `r CRANpkg("rmarkdown")` `r Citep(bib[["rmarkdown"]])` running behind the scenes. Citations were generated with `r CRANpkg("RefManageR")` `r Citep(bib[["RefManageR"]])`.

```{r vignetteBiblio, results="asis", echo=FALSE, warning=FALSE, message=FALSE}
## Print bibliography
PrintBibliography(bib, .opts = list(hyperlink = "to.doc", style = "html"))
```
