---
title: "Panel Catalogue"
author: 
  - name: Giulio Benedetti
    affiliation: University of Turku
    email: giulio.benedetti@utu.fi
output:
    BiocStyle::html_document:
        fig_height: 7
        fig_width: 10
        toc: true
        toc_float: true
        toc_depth: 2
        number_sections: true
        self_contained: true
        code_folding: show
date: "`r doc_date()`"
package: "`r pkg_ver('iSEEtree')`"
vignette: >
    %\VignetteIndexEntry{iSEEtree}
    %\VignetteEngine{knitr::rmarkdown}
    %\VignetteEncoding{UTF-8}
---

```{r, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Introduction

# Compositional Analysis

| Panel name             | Class name            | Plot function | Outline     |
|------------------------|-----------------------|---------------|-------------|
| Abundance plot         | AbundancePlot         | [plotAbundance](https://microbiome.github.io/miaViz/reference/plotAbundance.html) | Feature composition by sample |
| Abundance density plot | AbundanceDensityPlot  | [plotAbundanceDensity](https://microbiome.github.io/miaViz/reference/plotAbundanceDensity.html) | Feature distribution across samples |
| Prevalence plot        | PrevalencePlot        | [plotPrevalence](https://microbiome.github.io/miaViz/reference/plotPrevalence.html) | something |
| Complex heatmap plot   | ComplexHeatmapPlot    | fun | something |

## Abundance plot

## Abundance density plot

## Prevalence plot

## Complex heatmap plot

# Ordination Analysis

| Panel name             | Class name           | Plot function | Outline     |
|------------------------|----------------------|---------------|-------------|
| RDA plot               | RDAPlot              | [plotRDA](https://microbiome.github.io/miaViz/reference/plotCCA.html) | supervised ordination |
| Scree plot             | ScreePlot            | [plotScree](https://microbiome.github.io/miaViz/reference/plotScree.html) | Explained variance by component  |
| Loading plot           | LoadingPlot          | [plotLoadings](https://microbiome.github.io/miaViz/reference/plotLoadings.html) | Feature loadings by component |
| Reduced dimension plot | ReducedDimensionPlot | fun | any ordination result |

## RDA plot

## Scree plot

## Loading plot

## Reduced dimension plot

# Structural Analysis

| Panel name         | Class name       | Plot function | Outline     |
|--------------------|------------------|---------------|-------------|
| Row tree plot      | RowTreePlot      | [plotRowTree](https://microbiome.github.io/miaViz/reference/plotTree.html) | hierarchical structure of features |
| Column tree plot   | ColumnTreePlot   | [plotColTree](https://microbiome.github.io/miaViz/reference/plotTree.html) | hierarchical structure of samples  |
| Row graph plot     | RowGraphPlot     | [plotRowGraph](https://microbiome.github.io/miaViz/reference/plotGraph.html) | network structure of features |
| Column graph plot  | ColumnGraphPlot  | [plotColGraph](https://microbiome.github.io/miaViz/reference/plotGraph.html) | network structure of samples |

## Row/Column tree plots

## Row/Column graph plots

# Other panels

| Panel name         | Class name       | Plot function | Outline     |
|--------------------|------------------|---------------|-------------|
| Row tile plot      | RowTreePlot      | [plotRowTile](https://microbiome.github.io/miaViz/reference/plotColTile.html) | variable distribution across feature groups |
| Column tile plot   | ColumnTreePlot   | [plotColTile](https://microbiome.github.io/miaViz/reference/plotColTile.html) | variable distribution across sample groups  |
| Mediation plot     | MediationPlot    | [plotMediation](https://microbiome.github.io/miaViz/reference/plotMediation.html) | results of mediation analysis |
| Row data plot  | RowDataPlot  | fun | variables from feature metadata |
| Column data plot  | ColumnDataPlot  | fun | variables from sample metadata |

## Row/Column tile plots

## Mediation plot

## Row/Column data plots

# Reproducibility

R session information:

```{r reproduce, echo=FALSE}
## Session info
options(width = 120)
sessionInfo()
```

# References