---
title: "Panel Catalogue"
author: 
  - name: Giulio Benedetti
    affiliation: University of Turku
    email: giulio.benedetti@utu.fi
output:
    BiocStyle::html_document:
        fig_height: 7
        fig_width: 10
        toc: true
        toc_float: true
        toc_depth: 2
        number_sections: true
        self_contained: true
        code_folding: show
date: "`r doc_date()`"
package: "`r pkg_ver('iSEEtree')`"
vignette: >
    %\VignetteIndexEntry{iSEEtree}
    %\VignetteEngine{knitr::rmarkdown}
    %\VignetteEncoding{UTF-8}
---

```{r, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Introduction

This page introduces users to the complete catalogue of panels provided by
iSEEtree. Each panel is presented individually and visualised as it appears in
the app. This catalogue is divided into four sections:

- [compositional analysis](#sec-composition): abundance plot, abundance density
  plot, prevalence plot and complex heatmap plot
- [ordination analysis](#sec-ordination): RDA plot, Scree plot, loading plot and
  reduced dimension plot
- [structural analysis](#sec-structure): row/column tree plots and row/column
  graph plots
- [other](#sec-other): row/column tile plots, mediation plot and row/column data
  plots

# Compositional Analysis {#sec-composition}

| Panel name             | Panel class          | Purpose                             |
|------------------------|----------------------|-------------------------------------|
| Abundance plot         | AbundancePlot        | Feature composition by sample       |
| Abundance density plot | AbundanceDensityPlot | Feature distribution across samples |
| Prevalence plot        | PrevalencePlot       | Feature prevalence across samples   |
| Complex heatmap plot   | ComplexHeatmapPlot   | Whole assay composition             |

## Abundance plot

## Abundance density plot

## Prevalence plot

## Complex heatmap plot

# Ordination Analysis {#sec-ordination}

| Panel name             | Panel class          | Purpose                         |
|------------------------|----------------------|---------------------------------|
| RDA plot               | RDAPlot              | Supervised ordination           |
| Scree plot             | ScreePlot            | Explained variance by component |
| Loading plot           | LoadingPlot          | Feature loadings by component   |
| Reduced dimension plot | ReducedDimensionPlot | Any ordination result           |

## RDA plot

Supported operations:

- customising aesthetics

## Scree plot

## Loading plot



## Reduced dimension plot

# Structural Analysis {#sec-structure}

| Panel name         | Panel class      | Purpose                            |
|--------------------|------------------|------------------------------------|
| Row tree plot      | RowTreePlot      | Hierarchical structure of features |
| Column tree plot   | ColumnTreePlot   | Hierarchical structure of samples  |
| Row graph plot     | RowGraphPlot     | Network structure of features      |
| Column graph plot  | ColumnGraphPlot  | Network structure of samples       |

## Row/Column tree plots

[RowTreePlot]() and [ColumnTreePlot]() belong to the [TreePlot]() panel family.

Supported operations:

- collapsing and expanding clades
- rotating and opening trees
- ordering trees
- labeling nodes and tips
- selecting tree layout
- customising aesthetics

## Row/Column graph plots

# Other panels {#sec-other}

| Panel name         | Panel class    | Purpose                                     |
|--------------------|----------------|---------------------------------------------|
| Row tile plot      | RowTreePlot    | Variable distribution across feature groups |
| Column tile plot   | ColumnTreePlot | Variable distribution across sample groups  |
| Mediation plot     | MediationPlot  | Results of mediation analysis               |
| Row data plot      | RowDataPlot    | Variables from feature metadata             |
| Column data plot   | ColumnDataPlot | Variables from sample metadata              |

## Row/Column tile plots

## Mediation plot

## Row/Column data plots

# Reproducibility

R session information:

```{r reproduce, echo=FALSE}
## Session info
options(width = 120)
sessionInfo()
```

# References